We are not the 99 percent: quantifying asphericity in the distribution of Local Group satellites

### 2018-02-08 paper/paper.tex
edit

The aim of our work is to put together some of these pieces in a 
simple framework to quantify how atypical is the satellite
distribution found in the Local Group.
Some of the explicit characteristics we want to keep in our treatment
of the problem are the (i) cosmological context, (ii) the LG as a pair
of galaxies and (iii) a simple and robust description for the
satellites.

The first condition constrains the kind of simulations we use. 
We either results from fully cosmological simulations or resimulations
of cosmological sub-volumes. 
The second condition constraints the kind of samples we want to
use. We build explicit samples of pairs that resemble the kinematic
structure of the MW and M31. 
The third condition motivates to only use the inertia tensor as a
description for the satellites positions, while droping any
information that might come from 3D velocities (high uncertainties) or
involved algorithms for plane fitting.

Finally, to compute the odds of finding the LG satellite distributions 
in LCDM simulations, we do not make a direct search for the 
observed values into simulations because, as we show in Section, the
LG is itself difficult to fina and we do not have enough simulated systems to
perform such a brute force solution. 
We also avoid this direct comparison because we do not want to
implicitly use either LCDM or observations as what is considered
normal or standard and compute the deviations from that point.

Our approach is different. We set as a point of reference an spherical satellite
distribution to quantify how both simulations and observations
deviate from the spherical distribution, following recent
recommendations presented in \citep{2017AN....338..854P}

Simulations help us to construct an explicit probability distribution
for deviations from sphericity in LCDM.
Then we compute the odds of finding a satellite distribution that
deviates from sphericity as much as the LG does.

>>>>

The aim of our work is to overcome two difficulties in the study of
atipicallity in satellite distributions.
First, atypicallity is usually computed from distributions comming from
different simulated pair samples, making it inconvenient to compare
results from different studies. 
Second, halo pairs with similar to the LG are scarse in cosmological
volumes, making it hard to make direct comparisons against simulated
systems that fully resemble the LG.

We surmount the first difficulty by setting as a point of reference an
spherical satellite distribution and not LCDM simulations. 
This spherical distribution is built from the data itself
(observational or simulated) by randomizing its angular position
around the central galaxy and keeping fixed its radial distance
\citep{2017AN....338..854P}.

We handle the second difficuly by building an analytic probability
distribution that is compatible with the simulated data.
We the use the simulations to estimate the parameters in this
distribution to later use it as a parent sample to generate any
desired number of samples that are by construction statistically
compatible with simulations.

To summarize, simulations help us to construct an explicit probability
distribution for deviations from sphericity for a given LCDM simulated
sample.  
Using this distribution we compute the odds of finding a satellite
distribution that deviates from sphericity as much as the LG does.
